---
title: "Robust Support Vector Machines"
author: Zhu Wang^[UT Health San Antonio, zhuwang@gmail.com]
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: mpath.bib
link-citations: true
output:
  pdf_document: default
vignette: >
   %\VignetteIndexEntry{Robust Support Vector Machines}
   %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #tidy.opts=list(width.cutoff=60),tidy=TRUE)
```
The CC-family contains functions of composite of concave and convex functions. The CC-estimators are derived from minimizing loss functions in the CC-family by the iteratively reweighted convex optimization (IRCO), an extension of the iteratively reweighted least squares (IRLS). The IRCO reduces the weight of the observation that leads to a large loss; it also provides weights to help identify outliers. The CC-estimators include robust support vector machines. See @wang2020unified.

## Support vector machine classification

```{r, message=FALSE}
library("mpath")
library("e1071")
set.seed(1900)
x <- matrix(rnorm(40*2), ncol=2)
y <- c(rep(-1, 20), rep(1, 20))
x[y==1,] <- x[y==1, ] + 1
plot(x[,2],x[,1], col=(2-y))
```

Use the radial kernel SVM for classification.
```{r}
dat <- data.frame(x=x, y=as.factor(y))
svm.model <- svm(y~., data=dat, cost=100, type="C-classification")
summary(svm.model)
plot(svm.model, dat)
```

Robust radial kernel SVM for classification.
```{r, eval=TRUE}
ccsvm.model <- ccsvm(y ~ ., data = dat, cost = 100, type="C-classification", cfun="acave", 
                     s=1)
summary(ccsvm.model)
plot(ccsvm.model, dat)
```

Add 15% outliers to the training data, and fit robust SVM, selecting tuning parameters with the cross-validation method.
```{r, eval=TRUE}
n <- length(y)
nout <- n*0.15
id <- sample(n)[1:nout]
cat("id=", id)
y[id] <- -y[id]
dat2 <- data.frame(x=x, y=as.factor(y))
ccsvm.opt <- cv.ccsvm(y ~ ., data=dat2, type="C-classification", s=1, cfun="acave", 
                      n.cores=2, balance=FALSE)
ccsvm.opt$cost
ccsvm.opt$gamma
ccsvm.opt$s
```

To evaluate prediction, we simulate test data with no outliers.
```{r}
xtest <- matrix(rnorm(20*2), ncol=2)
ytest <- sample(c(-1,1), 20, rep=TRUE)
xtest[ytest==1, ] <- xtest[ytest==1, ] + 1
testdat <- data.frame(x=xtest, y=as.factor(ytest))
```

Fit a robust SVM model again, with tuning parameters selected by cross-validation, then evaluate prediction accuracy with test data, with 85% accuracy.
```{r, eval=TRUE}
ccsvm.model1 <- ccsvm(y ~ ., data = dat2, cost = ccsvm.opt$cost, gamma=ccsvm.opt$gamma, 
                      s=ccsvm.opt$s, cfun="acave", type="C-classification")
summary(ccsvm.model1)
table(predict=predict(ccsvm.model1, xtest), truth=testdat$y)
plot(ccsvm.model1, dat2)
```

Develop a SVM model with training data and evaluate with the test data. The prediction accuracy is 80%.
```{r, eval=TRUE}
svm.model1 <- svm(y~., data=dat2, cost=ccsvm.opt$cost, gamma=ccsvm.opt$gamma,  
                  type="C-classification")
summary(svm.model1)
table(predict=predict(svm.model1, testdat), truth=testdat$y)
plot(svm.model1, dat2)
```

In robust SVM with function \texttt{ccsvm}, argument \texttt{cfun} can be chosen from \texttt{"hcave", "acave", "bcave", "ccave", "dcave", "gcave", "tcave", "ecave"}, for a variety of concave functions.

## Support vector machine regression

We predict median value of owner-occupied homes in suburbs of Boston. The data can be obtained from the UCI machine learning data repository. There are 506 observations and 13 predictors.
```{r}
urlname <- "https://archive.ics.uci.edu/ml/"
filename <- "machine-learning-databases/housing/housing.data"
dat <- read.table(paste0(urlname, filename), sep="", header=FALSE)
n <- dim(dat)[1]
p <- dim(dat)[2]
cat("n=",n,"p=", p, "\n")
```
Randomly split the data into 90% of samples for training and 10% of samples as test data. 
```{r}
set.seed(129)
trid <- sample(n)[1:(n*0.9)]
traindat <- dat[trid, ]
testdat <- dat[-trid, ]
```
Fit the robust radial kernel CCSVM model with truncated $\epsilon$-insensitive loss, i.e., \texttt{cfun=``tcave''} in function \texttt{ccsvm}. Root mean squared error on test data is reported. A comprehensive robust CCSVM analysis  with other types of \texttt{cfun} can be found in @wang2020unified.
```{r}
ccsvm.model <- ccsvm(x=traindat[,-p], y=traindat[,p], cost = 2^3, gamma=2^(-4), 
                     epsilon=2^(-4), s=5, cfun="tcave")
summary(ccsvm.model)
ccsvm.predict <- predict(ccsvm.model, testdat[,-p])
mse1 <- mean((testdat[,p] - ccsvm.predict)^2)
cat("RMSE with robust SVM", sqrt(mse1))
```
Fit the radial kernel SVM model. The RMSE is larger than the robust SVM, and the model has a larger number of support vectors as well. See the figure below for a comparison.
```{r}
svm.model <- svm(x=traindat[,-p], y=traindat[,p], cost=2^3, gamma=2^(-4), epsilon=2^(-4))
summary(svm.model)
svm.predict <- predict(svm.model, testdat[,-p])
mse2 <- mean((testdat[,p] - svm.predict)^2)
cat("RMSE with SVM", sqrt(mse2))
plot(testdat[,p], ccsvm.predict, col="red", pch=1, ylab="Predicted values",
     xlab="Median home values ($1000)")
points(testdat[,p], svm.predict, col="black", pch=2)
legend("topleft", c("CCSVM", "SVM"), col=c("red", "black"), pch=c(1, 2))
abline(coef=c(0, 1))
```

## Reference
